---
output: github_document
---

# MyPackage

<!-- badges: start -->
<!-- badges: end -->

EnvEx provides functions to model non-stationary and joint extremes. Examples for a common metocean use case are provided.

## Installation

You can install the development version from GitHub with:

```r
```

## Example

This is a basic example showing storm and peak picking.

```{r message=FALSE, warning=FALSE}
# If installed:
library(evgam)
library(envex)

# subset ds_ekofisk to what is needed

ds <- df_ekofisk[, c("time", "hs", "wind_speed_10m", "tp", "tm2", "Pdir", "doy", "dt")]

# reduce data for test from 1976-01-01 to 1986-01-01
# ds_sub <- ds[1:87673,]  # 10 yrs for quicker testing

# choose primary variable
prime_varstr = 'hs'

# make formula
nknots <- c(4, 6)  # c(8, 12), c(6, 8) is default c(4, 6) for testing
formula_string <- paste(prime_varstr, "~ te(doy, Pdir, bs = c('cc', 'cc'), k = nknots)")
peak_picking_thr_model_fml <- as.formula(formula_string)

# decorrelation time scale of 1 day
df_pick <- peak_picking(ds, peak_picking_thr_model_fml,
                        decorrelation_time_scale = 1)

# additional variables were added, i.e. exc, thr, and storm_idx
```

## Plot the resulting storms, storm peaks, and thresholds

```{r fig.width=9, fig.height=4}
visualize_storm_picking(dfin = df_pick, dfinall = ds,
                        xstr = "dt", ystr = "hs",
                        sidx = 1, eidx = 4200)
```

## Plot the Hs density

```{r fig.width=9, fig.height=4}
par(mfrow = c(1, 3))
plot(density(df_pick$pots$hs, bw=.1))
plot(density(df_pick$storms$hs, bw=.1))
plot(density(ds$hs, bw=.1))
```

```{r message=FALSE, warning=FALSE}
## Model the non-stationary storm peaks

# choose primary variable, should be same as in example_picking.R script
prime_varstr = 'hs'

# number of bootstrap/resampling steps
nbstrp <- 100  # e.g. nbstrp = 10 for testing

system.time(res_bstrp <- run_bootstrap_storms_pots(df = df_pick$storms,
                                                   group_col = "storm_idx",
                                                   n_boot = nbstrp,
                                                   max_var = prime_varstr))
bstrp_storms_lst <- res_bstrp$boot_samples
bstrp_pots_lst <- res_bstrp$boot_max

for (b in seq_len(nbstrp-1)) {
  plot(density(bstrp_storms_lst[[b]]$hs, bw=.1), xlim=c(0, 15), ylim=c(0., .5), xlab = "", ylab = "", xaxt = "n", yaxt = "n", main = "")
  par(new=TRUE)
}
plot(density(bstrp_storms_lst[[b+1]]$hs, bw=.1), xlim=c(0,15), ylim=c(0.,.5), main = "", xlab = "Hs", ylab = "Density")

for (b in seq_len(nbstrp-1)) {
  plot(density(bstrp_pots_lst[[b]]$hs, bw=.1), xlim=c(0,15), ylim=c(0.,.5), xlab = "", ylab = "", xaxt = "n", yaxt = "n", main = "")
  par(new=TRUE)
}
plot(density(bstrp_pots_lst[[b+1]]$hs, bw=.1), xlim=c(0,15), ylim=c(0.,.5), main = "", xlab = "Hs", ylab = "Density")

```

## Notes

## License

## Development

To rebuild this README:

```r
devtools::build_readme()
```
