---
output: github_document
---

# MyPackage

<!-- badges: start -->
<!-- badges: end -->

EnvEx provides functions to model non-stationary and joint extremes. Examples for a common metocean use case are provided.

## Installation

You can install the development version from GitHub with:

```r
```

## Example

This is a basic example showing storm and peak picking.

```{r message=FALSE, warning=FALSE}
# If installed:
library(evgam)
library(envex)

# subset ds_ekofisk to what is needed

ds <- df_ekofisk[, c("time", "hs", "wind_speed_10m", "tp", "tm2", "Pdir", "doy", "dt")]

# reduce data for test from 1976-01-01 to 1986-01-01
# ds[1:87673,]  # 10 yrs for quicker testing

# choose primary variable
prime_varstr = 'hs'

# make formula
nknots <- c(4, 6)  # c(8, 12), c(6, 8) is default c(4, 6) for testing
formula_string <- paste(prime_varstr, "~ te(doy, Pdir, bs = c('cc', 'cc'), k = nknots)")
peak_picking_thr_model_fml <- as.formula(formula_string)

# decorrelation time scale of 1 day
df_pick <- peak_picking(ds, peak_picking_thr_model_fml,
                        decorrelation_time_scale = 1)

# additional variables were added, i.e. exc, thr, and storm_idx
```

## Plot the resulting storms, storm peaks, and thresholds

```{r fig.width=9, fig.height=4}
visualize_storm_picking(dfin = df_pick, dfinall = ds,
                        xstr = "dt", ystr = "hs",
                        sidx = 1, eidx = 4200)
```

## Plot the Hs density

```{r fig.width=9, fig.height=4}
par(mfrow = c(1, 3))
plot(density(df_pick$pots$hs, bw=.1))
plot(density(df_pick$storms$hs, bw=.1))
plot(density(ds$hs, bw=.1))
```

## Bootstrap the storms

```{r message=FALSE, warning=FALSE}
# number of bootstrap/resampling steps
nbstrp <- 20  # e.g. nbstrp = 20 for testing

system.time(res_bstrp <- run_bootstrap_storms_pots(df = df_pick$storms,
                                                   group_col = "storm_idx",
                                                   n_boot = nbstrp,
                                                   max_var = prime_varstr))
bstrp_storms_lst <- res_bstrp$boot_samples
bstrp_pots_lst <- res_bstrp$boot_max
```
## Plot the sampled within storm Hs densities

```{r fig.width=9, fig.height=4}
for (b in seq_len(nbstrp-1)) {
  plot(density(bstrp_storms_lst[[b]]$hs, bw=.1), xlim=c(0, 15), ylim=c(0., .5), xlab = "", ylab = "", xaxt = "n", yaxt = "n", main = "")
  par(new=TRUE)
}
plot(density(bstrp_storms_lst[[b+1]]$hs, bw=.1), xlim=c(0,15), ylim=c(0.,.5), main = "", xlab = "Hs", ylab = "Density")
```

## Plot the sampled storm peak Hs densities

```{r fig.width=9, fig.height=4}
## Plot the within storm Hs
for (b in seq_len(nbstrp-1)) {
  plot(density(bstrp_pots_lst[[b]]$hs, bw=.1), xlim=c(0,15), ylim=c(0.,.5), xlab = "", ylab = "", xaxt = "n", yaxt = "n", main = "")
  par(new=TRUE)
}
plot(density(bstrp_pots_lst[[b+1]]$hs, bw=.1), xlim=c(0,15), ylim=c(0.,.5), main = "", xlab = "Hs", ylab = "Density")
```

## Model the non-stationary storm peaks

```{r echo=TRUE, message=FALSE, warning=FALSE, results='hide'}
library(ppgam)
library(extraDistr)
library(ggplot2)
library(extRemes)
# define knots for cyclic splines
knots = list(doy = c(0,366), Pdir=c(0,360))
# indicate variable that needs ppgam weights
node_str_lst = c('Pdir')
# number of years covered by original dataset
nr_of_years <- 48

# threshold range from cross-validation
thr_range <- list('hs'=c(.8,.9))

# define models
fml_pp <- ~ te(doy, Pdir, bs = c('cc', 'cc'), k = c(6,8))
fml_gpd <- list(exc ~ te(doy, Pdir, k=c(6,8), bs=c("cc","cc")),
                ~ te(doy, Pdir, k=c(6,8), bs=c("cc","cc")))
fml_ald <- hs ~ te(doy, Pdir, bs=c("cc","cc"), k=c(6,8))
fml_ald_tm2 <- tm2 ~ te(doy, Pdir, bs=c("cc","cc"), k=c(6,8))

model_fml_thr <- NULL
model_fml_occ <- NULL
model_fml_gpd <- NULL
model_fml_thr[['hs']] <- fml_ald
model_fml_occ[['hs']] <- fml_pp
model_fml_gpd[['hs']] <- fml_gpd
model_fml_thr[['tm2']] <- fml_ald_tm2
model_fml_occ[['tm2']] <- fml_pp
model_fml_gpd[['tm2']] <- fml_gpd


models_margs <- fit_margs_bstrp(bstrp_pots_lst,
                                model_fml_thr, model_fml_occ, model_fml_gpd,
                                extr_thr=thr_range, nr_of_years,
                                thr_str = 'thr',
                                list_var = c('hs','tm2'),
                                nbstrp = nbstrp,
                                nquad = 225,
                                knots = knots,
                                node_str_lst = node_str_lst)
```

## Simulate 100yr maxima

```{r echo=TRUE, message=FALSE, warning=FALSE, results='hide'}
RP <- 100
nmc <- 100
preds_margs_lst <- predict_margs(models_margs, nr_yrs_subdata, RP=RP,
                                 nmc = nmc, var_lst = c('hs','tm2'),
				 nbstrp = nbstrp,
                                 covarstr_lst = c('Pdir','doy'),
                                 grid_interval = list('Pdir'=1, 'doy'=1))
```

## Plot resulting simulated peaks

```{r fig.width=9, fig.height=4}
dhs <- diagnose_margs_preds_density(preds_margs_lst, 'hs', bw=.1, xlim=c(5,40))
abline(v = max(df_pick$pots$hs), col = "gray", lwd = 2)
rvs <- NULL
for (i in 1:length(dhs$maxvals)){
  rvs[[i]] <- quantile(unlist(dhs$maxvals[[i]]), exp(-1))
}
rv100_q4 <- quantile(unlist(dhs$maxvals),exp(-1))  # q4 estimator
rv100_q2p <- mean(unlist(rvs))  # q'2 estimator
abline(v = rv100_q2p, col = "blue", lwd = 2)
abline(v = rv100_q4, col = "paleturquoise2", lwd = 2)
```

## Compute joint models testing various thresholds

```{r echo=TRUE, message=FALSE, warning=FALSE, results='hide'}
# check various thresholds ###
maxd_thr <- seq(.5,.98,.01)
models_maxds_thr <- NULL
for (t in seq_along(maxd_thr)){
 print(c('## t:', maxd_thr[t]))
 models_maxds_tmp <- fit_maxds_bstrp(bstrp_pots_lst,
                                     models_margs,
                                     maxd_thr[t],
                                     nbstrp=nbstrp)
 models_maxds_thr[[t]] <- models_maxds_tmp
}
```

```{r fig.width=9, fig.height=4}
# diagnose effect of threshold choice
diagnose_maxds(models_maxds_thr, maxd_thr, 'hs', 'tm2', ylim=c(-1.5, 1.5))
```

## Choose threshold range from test and sample across

```{r echo=TRUE, message=FALSE, warning=FALSE, results='hide'}
maxd_thr_range <- c(.6,.8)
models_maxds <- fit_maxds_bstrp(bstrp_pots_lst,
                                models_margs,
                                maxd_thr,
                                nbstrp = nbstrp)
```

# Diagnose by comparing data against nsim HT2004 simulations

```{r fig.width=9, fig.height=4}
nsim <- 200
diagnose_maxds_fitted(models_maxds[[2]], "tm2", "hs", nsim)
```
```{r fig.width=9, fig.height=4}
diagnose_maxds_fitted(models_maxds[[2]], "hs", "tm2", nsim)
```

## Simulate joint 100yr events

```{r echo=TRUE, message=FALSE, warning=FALSE, results='hide'}
# predict on Laplace space
preds_maxds_LP <- predict_from_HT2004_models(models_margs, models_maxds,
                                             c('tm2','hs'), nbstrp,
                                             preds=preds_margs_lst)
# convert to real space
preds_maxds <- convert_HT2004_preds_to_original_space(models_maxds,
                                                      preds_maxds_LP,
						      preds_margs_lst,
                                                      var_lst = c("tm2", "hs"))

# retrieve valid predicted values according to chosen space
valids <- retrieve_valid_HT_samples(preds_maxds,
                                    preds_maxds_LP,
                                    preds_margs_lst,
                                    "hs", c("tm2"),
                                    models_maxds = models_maxds)


df_joint <- data.frame(unlist(valids$X_valid$tm2), unlist(valids$Y_valid$tm2))
colnames(df_joint) <- c("hs", "tm2")

dfcovar <- NULL
for (n in names(valids$covar_valid$tm2[[1]])){
  dfcovar[[n]] <- NULL
  for (b in 1:nbstrp){
    dfcovar[[n]][[b]] <- valids$covar_valid$tm2[[b]][[n]]
  }
}

# add covariates
df_joint[['doy']] <- unlist(dfcovar[['doy']])
df_joint[['Pdir']] <- unlist(dfcovar[['Pdir']])
```

## Plot Joint Distr

```{r fig.width=9, fig.height=4}
library(ggplot2)
library(metR)
specific_levels <- c(0.01,0.05,0.1)
ggplot() +
  geom_bin2d(data = df_joint,  aes(x = df_joint$hs, y = df_joint$tm2), bins = 50) +
  scale_fill_gradient(low = "white", high = "blue") +  # Color gradient
  geom_density_2d(data = df_joint,  aes(x = df_joint$hs, y = df_joint$tm2), color = "red", size = 0.1, breaks = specific_levels) +  # Density lines
  xlim(10, 21) +  # Extent in the x direction
  ylim(5, 14) +  # Extent in the y direction
  labs(title = "HT2004 predictions vs true annual joint maxima",
       x = "Hs [m]",
       y = "Tm2 [s]",
       fill = "Bin density") +
  theme_minimal()
```

## Simulate storms (matching) given simulated storm peaks

```{r echo=TRUE, message=FALSE, warning=FALSE, results='hide'}
df_hist_pots <- df_pick$pots
df_hist_storms <- df_pick$storms

# add steepness
df_joint_pots <- df_joint
df_hist_pots[['s_tm2']] <- compute_steepness(df_hist_pots$hs, df_hist_pots$tm2)
df_joint_pots[['s_tm2']] <- compute_steepness(df_joint_pots$hs, df_joint_pots$tm2)

dist_fct_lst <- NULL
dist_fct_lst[['Pdir']] <- 'circ'
dist_fct_lst[['doy']] <- 'circ'
dist_fct_lst[['hs']] <- 'euc'
dist_fct_lst[['tm2']] <- 'euc'
dist_fct_lst[['s_tm2']] <- 'euc'
dist_circ_L_lst <- NULL
dist_circ_L_lst[['Pdir']] <- 360
dist_circ_L_lst[['doy']] <- 365
varlst_matching <- c("hs","s_tm2","doy","Pdir")

nr_of_storms <- 100

df_sim_in <- df_joint_pots[varlst_matching][1:nr_of_storms,]
df_hist_in <- df_hist_pots[c("hs", "s_tm2", "doy" , "Pdir", "tm2", "storm_idx")]
matched_storm_idx <- find_idx_closest_storm_peaks(df_sim_in, df_hist_in,
                                                  varlst = varlst_matching,
                                                  dist_fct_lst = dist_fct_lst,
                                                  dist_circ_L_lst = dist_circ_L_lst,
                                                  sidx = 1, eidx = 100)

df_storms_filtered <- df_pick$storms[df_pick$storms$storm_idx %in% matched_storm_idx[1:nr_of_storms], ]
res <- anchor_sim_storms(df_joint_pots, df_hist_pots, df_hist_storms, matched_storm_idx)
```

```{r fig.width=9, fig.height=9}
vis_sim_storms(res, storm_idx=10, xlim=c(0,13), ylim=c(0,20))
```

```{r fig.width=9, fig.height=5}
show_sim_pop(res, xlim=c(0,20), ylim=c(0,40))
```


## Notes

## License

## Development

To rebuild this README:

```r
devtools::build_readme()
```
